Versioned data: why it is needed and how it can be achieved (easily and cheaply)

### 2017-11-09 ms.tex
final two paragraphs

While our first attempt \textsc{vdd} system, \texttt{datastorr} is currently being used, there remain several challenges that we anticipate will be overcome as both technology and scientific norms change. For example, because \texttt{git} was built for tracking changes in code, there are aspects of how it behaves that are not natural to apply to data \citep{Perkel-2016}. For example, because it tracks changes line-by-line rather than cell-by-cell, there are inefficiencies in how the software stores the changes. There is rapid current work on new implementations which will be faster computationally, smaller in file size, flexible with respect to data structures, and thus allow for larger datasets to be versioned efficiently \citep{Fli, Dat}. Another area requiring attention is properly allocating credit to authors for these citations, and we hope that the large citation mapping services (\texttt{Web of Science} and \texttt{Google Scholar}) will adapt to the community standard once the research community settles on best archiving practices for versioned data. Likewise, we are optimistic that our funding bodies, research institutions, and scientific colleagues will further recognize the value in adding to, improving, and curating existing datasets in addition to collecting new ones.

@@

>>>>

While our first attempt \textsc{vdd} system, \texttt{datastorr} is currently being used, there remain several challenges that we anticipate will be overcome as both technology and scientific norms change. Another area requiring attention is properly allocating credit to authors for these citations, and we hope that the large citation mapping services (\texttt{Web of Science} and \texttt{Google Scholar}) will adapt to the community standard once the research community settles on best archiving practices for versioned data. Likewise, we are optimistic that our funding bodies, research institutions, and scientific colleagues will further recognize the value in adding to, improving, and curating existing datasets in addition to collecting new ones.

@@

\noindent\makebox[\linewidth]{\rule{\paperwidth}{0.4pt}}

The design choices we made in datastorr represent only one of many possible ways to adopt a \textsc{vdd} system; we can imagine many alternative models for interacting with versioned data stored in the cloud. The key is not the specific technology, but rather the concept of creating, maintaining, and distributing versioned data.  Indeed, as with every technology the best available approach is certain to change: there are currently a number of promising technologies---CKAN, rdrop2 and Git-LFS---that may facilitate even better \textsc{vdd} systems in the future.  There is also rapid current work on new implementations which will make versioning systems faster computationally, smaller in file size, flexible with respect to data structures, and thus allow for larger datasets to be versioned efficiently \citep{Fli, Dat}.

In short, many of the key roadblocks preventing a switch from a static to a dynamic data world were technological: in the past, it took great deal of money and expertise to set up and maintain a \textsc{vdd} system, but now there are many platforms for storing and distributing versioned controlled data that are cheap and relatively easy to set up.
